---
title: "Introduction to Matrix Algebra"
author: "William Murrah"
date: ''
output:
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
bibliography: ref.bib
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE, }
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(
  comment = NULL,    
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)                
```

## Why should I learn matrix algebra?

For at least two reasons, one 

> [Matrix algebra (MA)] furnishes a compact language that unifies important aspects of univariate and mutivariate statistics. In this way, MA allows one to understand mutivariable techniques using their univariate analogues and counterparts [@raykov2011introduction, p. 21].


General notation for a matrix:

\[ \textbf{A} = \left[ \begin{array}{cccccc}
a_{11} & a_{12} & ... & a_{1p}\\
a_{21} & a_{22} & ... & a_{2p}\\
 .     & .      & .  & . \\
a_{n1} & a_{n2} & ... & a_{np}\end{array} \right],\] 

which is referred to as a $n\times p$ matrix.

A vector is a special kind of matrix used often in statistics. a row vector is a matrix with a single row and two or more columns, while a column vector is a vector with one column and two or more rows:

row vector:

\[x = \left[ \begin{array}{ccc}1, 3, 5,7\end{array} \right],\]

column vector:

\[ y = \left[\begin{array}{ccc}
1\\ 3\\ 5\\7 \end{array}\right].\]

Note that $x$ is a $1 \times 4$ row vector and $y$ is a $4 \times 1$ column matrix. The convention is that when the type of vector is not specified, it is a column vector. Think of variables in a data frame as iconic vectors, which are column vectors.

A scalar can be thought of as a one dimensional matrix.

## Matrix Operations

Two matrices are equal if:
  a. they have the same dimensions
  b. each element in one is identical to the element in the same position in the other.
  
\[ \textbf{C} = \left[ \begin{array}{cc}
2 & 3\\
3 & 2\\
4 & 2
\end{array} \right] \neq 
%
\left[ \begin{array}{cc}
2 & 2\\
3 & 3\\
4 & 2
\end{array} \right]  = \textbf{D},\]

\[ \textbf{E} = \left[ \begin{array}{cc}
1 & 4\\
2 & 5\\
3 & 6
\end{array} \right] \neq 
%
\left[ \begin{array}{cc}
4 & 1\\
5 & 2\\
6 & 3
\end{array} \right]  = \textbf{F},\]


\[ \textbf{G} = \left[ \begin{array}{cc}
1 & 4\\
2 & 5\\
3 & 6
\end{array} \right] \neq 
%
\left[ \begin{array}{ccc}
1 & 2 & 3\\
4 & 5 & 6
\end{array} \right]  = \textbf{H}.\]

```{r echo=FALSE}
sessionInfo()  
```
  
# References
  