---
title: "Measurement Model Scaling and Identification"
author: "William Murrah"
date: ''
output:
  html_document:
    fig_height: 3
    fig_width: 5
    keep_md: yes
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir ="../../../../")
opts_chunk$set(
  echo = FALSE,
  comment = NULL,
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)  
require(lavaan)
require(ztable)
options(ztable.type = 'html', ztable.zebra.color="platinum")
library(DiagrammeR)
```


## Conventional Drawing and Labeling of SEM Model



### SEM nomenclature conventions

| name   | upper case   | lower case   | usage                                                            |
|--------|----------------|------------|------------------------------------------------------------------|
| Lambda |$\Lambda$     | $\lambda$    | Loading of a manifest indicator onto a latent construct          |
| Psi    | $\Psi$       | $\psi$       | residual variance/covariance of contruct when endogenous         |
| Theta  | $\Theta$     | $\theta$     | residual variance/covariance of indicators                       |
| Sigma  | $\Sigma$     | $\sigma$     | $\Sigma$ is the model implied variance/covariance matrix;        |
|        |              |              | $\sigma$ is standard deviation, $\sigma^2$ variance of indicator.|
|        |              |              | $\sigma$ can also be covariance of indicator                     |


```{r}
grViz("
digraph Basic {
node [shape = circle]
Construct;
node [shape = box]
indicator_1; indicator_2; indicator_3;
# Edges
Construct -> indicator_1 [label = <&nbsp;&lambda;<sub>1,1</sub>>];
Construct -> indicator_2 [label = <&nbsp;&lambda;<sub>2,1</sub>>];
Construct -> indicator_3 [label = <&nbsp;&lambda;<sub>3,1</sub>>];
Construct:n -> Construct:n [dir=both, label = <&psi;<sub>1,1</sub>>];

indicator_1:s -> indicator_1:s [dir=both, label = <&theta;<sub>1,1</sub>>];
indicator_2:s -> indicator_2:s [dir=both, label = <&theta;<sub>2,2</sub>>];
indicator_3:s -> indicator_3:s [dir=both, label = <&theta;<sub>3,3</sub>>];

{rank = same; indicator_1; indicator_2; indicator_3;}
} 
")

```
\[ \textbf{$\Sigma$} = \left[ \begin{array}{cccccc}
\sigma_{1,1}^2 & \sigma_{1,2} & \sigma_{1,3} \\
\sigma_{2,1} & \sigma_{2,2}^2 & \sigma_{2,3} \\
\sigma_{3,1} & \sigma_{3,2} & \sigma_{3,3}^2 
\end{array} \right],\] 

\[ \textbf{$\Lambda$} = \left[ \begin{array}{cc}
\lambda_{1,1} \\
\lambda_{2,1} \\
\lambda_{3,1}
\end{array} \right],\] 

\[ \textbf{$\Psi$} = \left[ \begin{array}{cc}
\psi_{1,1} 
\end{array} \right],\] 

\[ \textbf{$\Lambda^\prime$} = \left[ \begin{array}{cc}
\lambda_{1,1} & \lambda_{2,1} & \lambda_{3,1} 
\end{array} \right],\] 

\[ \textbf{$\Theta$} = \left[ \begin{array}{cccccc}
\theta_{1,1} & 0 & 0  \\
0 & \theta_{2,2} & 0  \\
0 & 0 & \theta_{3,3} 
\end{array} \right].\] 

### Fundamental SEM equation

$$
\Sigma = \Lambda \Psi \Lambda' + \Theta \tag{1}
$$

## Measurement Model: 3 indicators


```{r}
dt <- grViz("
    
digraph CFA {

node [shape = circle]
Positive;

node [shape = box]
Glad; Cheer; Happy;

# Edges
Positive -> Glad [label = <&lambda;<sub>1</sub>>];
Positive -> Cheer [label = <&lambda;<sub>2</sub>>];
Positive -> Happy [label = <&lambda;<sub>3</sub>>];
Positive:n -> Positive:n [dir=both, label = <&psi;>,position = N]
Glad:s -> Glad:s [dir=both, label = <&theta;<sub>1</sub>>]
Cheer:s -> Cheer:s [dir=both, label = <&theta;<sub>2</sub>>]
Happy:s -> Happy:s [dir=both, label = <&theta;<sub>3</sub>>]

	{rank = same; Positive;}
	{rank = same; Glad; Cheer; Happy;}
}  
")
dt
```

```{r, echo = TRUE}
library(lavaan)

##Prepare data with sufficient statisitics##
mymeans<-matrix(c(3.06893, 2.92590, 3.11013), ncol=3,nrow=1)
mysd<-c(0.84194,0.88934,0.83470)
mat <- c(1.00000,
         0.55226, 1.00000,
         0.56256, 0.60307, 1.00000)
mycor <- getCov(mat, lower = TRUE)

##Transform correlation matrix to covariance matrix using information above##
mycov <- mysd %*% t(mysd)

rownames(mycor) <-c( "Glad", "Cheerful", "Happy")
colnames(mycor) <-c( "Glad", "Cheerful", "Happy")

rownames(mycov) <-c( "Glad", "Cheerful", "Happy")
colnames(mycov) <-c( "Glad", "Cheerful", "Happy")
mynob<-823
```


```{r}
sigma <- mycov
```


```{r, results = 'asis'}
cr <- ztable(mycor, zebra = 2)
tab <- data.frame(Mean = round(mymeans[1, ], 3), SD = round(mysd, 3), 
                  Var = round(mysd*mysd, 3))

pt <- t(cbind(mycor, tab))
ztable(pt, zebra = 2, caption = "Descriptive Statistics")

```

### Mplus input file (`marker1.inp`)

```{r}
file <-'topics/2_MeasurementModel/2b_ConfirmatoryFactorAnalysis/mplus/marker1.inp'
noquote(scan(file, what = 'character', sep = '\n'))
```


### Mplus output using Correlation Table and first indicator fixed to 1.00

```{r , }
file <-'topics/2_MeasurementModel/2b_ConfirmatoryFactorAnalysis/mplus/marker1.out'
noquote(scan(file, what = 'character', sep = '\n', skip = 112, nlines = 17))
```


```{r, echo=TRUE}
lambda = matrix(c(1.00, 1.072, 1.092), nrow = 3)
lambda
psi    = matrix(.515)
psi
t(lambda)
theta  = diag(c(.484, .408, .385))
theta
sigma = lambda %*% psi %*% t(lambda) + theta
round(sigma,2)
```

```{r, results='asis'}
ztable(mycor)
```

## Measurement Model: 6 indicators, 2 latent constructs

```{r}
grViz("
digraph Basic {
node [shape = circle]
Construct_1;
Construct_2;
node [shape = box]
indicator_1; indicator_2; indicator_3;
indicator_4; indicator_5; indicator_6;
# Edges
Construct_1 -> indicator_1 [label = <&nbsp;&lambda;<sub>1,1</sub>>];
Construct_1 -> indicator_2 [label = <&nbsp;&lambda;<sub>2,1</sub>>];
Construct_1 -> indicator_3 [label = <&nbsp;&lambda;<sub>3,1</sub>>];
Construct_1:n -> Construct_1:n [dir=both, label = <&psi;<sub>1,1</sub>>];

indicator_1:s -> indicator_1:s [dir=both, label = <&theta;<sub>1,1</sub>>];
indicator_2:s -> indicator_2:s [dir=both, label = <&theta;<sub>2,2</sub>>];
indicator_3:s -> indicator_3:s [dir=both, label = <&theta;<sub>3,3</sub>>];

Construct_2 -> indicator_4 [label = <&nbsp;&lambda;<sub>4,2</sub>>];
Construct_2 -> indicator_5 [label = <&nbsp;&lambda;<sub>5,2</sub>>];
Construct_2 -> indicator_6 [label = <&nbsp;&lambda;<sub>6,2</sub>>];
Construct_2:n -> Construct_2:n [dir=both, label = <&psi;<sub>2,2</sub>>];

indicator_4:s -> indicator_4:s [dir=both, label = <&theta;<sub>4,4</sub>>];
indicator_5:s -> indicator_5:s [dir=both, label = <&theta;<sub>5,5</sub>>];
indicator_6:s -> indicator_6:s [dir=both, label = <&theta;<sub>6,6</sub>>];

Construct_1:n -> Construct_2:n [dir=both, label = <&psi;<sub>2,1</sub>>];

{rank = same; Construct_1; Construct_2}
{rank = same; indicator_1; indicator_2; indicator_3; 
indicator_4; indicator_5; indicator_6;}
} 
")

```

\[ \textbf{$\Sigma$} = \left[ \begin{array}{cccccc}
\sigma_{1,1}^2 & \sigma_{1,2} & \sigma_{1,3} & \sigma_{1,4} & \sigma_{1,5} & \sigma_{1,6} \\
\sigma_{2,1} & \sigma_{2,2}^2 & \sigma_{2,3} & \sigma_{2,4} & \sigma_{2,5} & \sigma_{2,6} \\
\sigma_{3,1} & \sigma_{3,2} & \sigma_{3,3}^2 & \sigma_{3,4} & \sigma_{3,5} & \sigma_{3,6} \\
\sigma_{4,1} & \sigma_{4,2} & \sigma_{4,3} & \sigma_{4,4}^2 &  \sigma_{4,5} & \sigma_{4,6} \\
\sigma_{5,1} & \sigma_{5,2} & \sigma_{5,3} & \sigma_{5,4} & \sigma_{5,5}^2 & \sigma_{5,6} \\
\sigma_{6,1} & \sigma_{6,2} & \sigma_{6,3} & \sigma_{6,4} & \sigma_{6,5} & \sigma_{6,6}^2 
\end{array} \right],\] 

\[ \textbf{$\Lambda$} = \left[ \begin{array}{cc}
\lambda_{1,1} & 0 \\
\lambda_{2,1} & 0 \\
\lambda_{3,1} & 0 \\
0 & \lambda_{4,2} \\
0 & \lambda_{5,2} \\
0 & \lambda_{6,2}
\end{array} \right],\] 

\[ \textbf{$\Psi$} = \left[ \begin{array}{cc}
\psi_{1,1} & \psi_{1,2} \\
\psi_{2,1} & \psi_{2,2}
\end{array} \right],\] 

\[ \textbf{$\Lambda^\prime$} = \left[ \begin{array}{cc}
\lambda_{1,1} & \lambda_{2,1} & \lambda_{3,1} 
0 & 0 & 0\\
0 & 0 & 0 & \lambda_{4,2} & \lambda_{5,2} & \lambda_{6,2}
\end{array} \right],\] 

\[ \textbf{$\Theta$} = \left[ \begin{array}{cccccc}
\theta_{1,1} & 0 & 0 & 0 & 0 & 0 \\
0 & \theta_{2,2} & 0 & 0 & 0 & 0 \\
0 & 0 & \theta_{3,3} & 0 & 0 & 0 \\
0 & 0 & 0 & \theta_{4,4} &  0 & 0 \\
0 & 0 & 0 & 0 & \theta_{5,5} & 0 \\
0 & 0 & 0 & 0 & 0 & \theta_{6,6} 
\end{array} \right].\] 
```{r, echo = TRUE}
lambda = matrix(c(.712, .788, .768, 0.0, 0.0, 0.0,
                  0.0, 0.0, 0.0, .729, .764, .778), nrow = 6)
lambda
psi    = matrix(c(1.00, 0.561,
                  0.561, 1.00), nrow = 2)
psi
theta = diag(c(.491, .378, .409, .467, .416, .394))
theta

sigma = lambda %*% psi %*% t(lambda) + theta
round(sigma, 2)
```

```{r, echo = TRUE, results = 'asis'}

vec <- dput(scan('topics/2_MeasurementModel/2b_ConfirmatoryFactorAnalysis/mplus/PAcorr2t.dat', quiet = TRUE))
mat <- getCov(vec, lower = TRUE)
ztable(mat)
```


## Adding Means to the Model

#### The means structure equation:
$$
\mu_y = \textbf{T} + \Lambda \textbf{A}. \tag{2}
$$


```{r}
dt <- grViz("
    
digraph CFA {

node [shape = circle]
Positive;

node [shape = box]
Glad; Cheer; Happy;

node [shape = triangle]

1;
# Edges
Positive -> Glad [label = <&lambda;<sub>1</sub>>];
Positive -> Cheer [label = <&lambda;<sub>2</sub>>];
Positive -> Happy [label = <&lambda;<sub>3</sub>>];
Positive:n -> Positive:n [dir=both, label = <&psi;>,position = N]
Glad:s -> Glad:s [dir=both, label = <&theta;<sub>1</sub>>]
Cheer:s -> Cheer:s [dir=both, label = <&theta;<sub>2</sub>>]
Happy:s -> Happy:s [dir=both, label = <&theta;<sub>3</sub>>]

1-> Positive [label = 0];
1 -> Glad [label = <&tau;<sub>1</sub>>];
1 -> Cheer [label = <&tau;<sub>2</sub>>];
1 -> Happy [label = <&tau;<sub>3</sub>>];

{rank = same; Positive; 1;}
{rank = same; Glad; Cheer; Happy;}
}  
")
dt
```
```{r}
# grViz("
#     
# digraph CFA {
# 
# node [shape = circle]
# Positive;
# 
# node [shape = box]
# Glad; Cheer; Happy;
# 
# # Edges
# Positive -> Glad [label = <&nbsp;&lambda;<sub>1</sub>>];
# Positive -> Cheer [label = <&nbsp;&lambda;<sub>2</sub>>];
# Positive -> Happy [label = <&nbsp;&lambda;<sub>3</sub>>];
# Positive:n -> Positive:n [dir=both, label = <&psi;>]
# Glad:s -> Glad:s [dir=both, label = <&theta;<sub>1</sub>>]
# Cheer:s -> Cheer:s [dir=both, label = <&theta;<sub>2</sub>>]
# Happy:s -> Happy:s [dir=both, label = <&theta;<sub>3</sub>>]
# 
# 	{rank = same; Positive;}
# 	{rank = same; Glad; Cheer; Happy;}
# }  
# ")

```


```{r}
# mod <- '
# Positive =~ Glad + Cheerful + Happy
# '
# 
# fit <- cfa(mod, sample.cov=mycov, sample.nobs = mynob, sample.mean=mymeans, 
#           std.lv=T)
# summary(fit)

```


```{r}
# lambda <- matrix(c(1.00, 
#                    1.072,
#                    1.092), nrow=3)
# psi <- .515
# 
# theta <- diag(c(.484, .408, .385))
# 
# sigma <- lambda %*% psi %*% t(lambda) + theta
```





```{r}
# grViz("
# digraph Cheer {
# 
# node [shape = circle]
# Positive;
# 
# node [shape = box]
# Cheer;
# 
# # Edges
# 
# Positive -> Cheer [label = <&nbsp;&lambda;>];
# Positive:n -> Positive:n [dir=both, label = <&psi;>,position = N];
# Cheer:s -> Cheer:s [dir=both, label = <&theta;>];
# }  
# ")
```

```{r}
# mod1<-'Positive =~ 1*Cheerful
# Positive~~Positive
# Cheerful~~0*Cheerful'
# #Save output to fit1##
# fit1<-cfa(mod1, sample.cov=mycov, sample.nobs = mynob, sample.mean=mymeans, 
#           std.lv=F)

##Request for summary of output##


# summary(fit1, fit.measures=T)
```